---
title: "Data preparation"
author: "Stefano Coretta"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
theme_set(theme_light())
library(googlesheets4)
library(magrittr)
library(FactoMineR)
library(factoextra)
```

# Read data

We first read the Emilian and Esperanto questionnaire data. Tidy column names are added.

```{r read-data, message=FALSE}
# Emilian and Esperanto headings
headings_em <- read_csv("data/raw/headings-emilian.csv")
headings_eo <- read_csv("data/raw/headings-esperanto.csv")
space_type <- read_sheet("https://docs.google.com/spreadsheets/d/1hYs92kMU34zPi2Ibfj9PmqjtsnGJqvqQOeplZVg0JGQ/edit?usp=sharing", 3) %>%
  filter(
    !is.na(space_type)
  )

emilian <- read_csv("data/raw/emilian-clean.csv", skip = 1, col_names = headings_em$new, na = c("", "NA", "na"))

esperanto <- read_csv("data/raw/esperanto-clean.csv", skip = 1, col_names = headings_eo$new, na = c("", "NA", "na")) %>%
  mutate(
    # Clean age column
    age = str_remove(age, "jaroj"),
    age = str_replace(age, "naskiĝis en la 1995a", "25"),
    age = as.numeric(age)
  )
```

# Bind Emilian and Esperanto data

We can now bind the data together, apply recoding to some columns and create new columns.

```{r emilianto}
#### Main data table ####
emilianto <- bind_rows(
  emilian %>% mutate(language = "Emilian"),
  esperanto %>% mutate(language = "Esperanto")
) %>%
  
  #### Refactor and recode columns ####
  mutate(
    # Fix coding error MD+ > MD
    languages_family = ifelse(languages_family == "MD+", "MD", languages_family),
    # Refactor understand and speak
    understand = ordered(understand, levels = c("NO", "AL", "50/50", "G", "VG")),
    speak = ordered(speak, levels = c("NO", "AL", "50/50", "G", "VG")),
    across(educated:familiar, ~ as.ordered(.x)),
    # Recode columns
    profession_2 = case_when(
      profession == "STUDENT" ~ "student",
      profession %in% c("RETIRED", "NOTIW") ~ "not in work",
      profession %in% c("ISCO-AGR", "ISCO-CLE", "ISCO-SER", "ISCO-LAB", "ISCO-ELE") ~ "unskilled",
      profession %in% c("ISCO-MAN", "ISCO-PRO", "ISCO-TEC", "ISCO-CRA", "ISCO-ARM") ~ "skilled"
    ),
    languages_family_2 = case_when(
      languages_family %in% c("BD", "MULTD", "BE", "MULTE") ~ "multi_e",
      languages_family %in% c("BO", "MULTO") ~ "multi_o",
      languages_family %in% c("MD", "ME") ~ "mono_e",
      languages_family %in% c("MI", "MO") ~ "mono_o",
    ),
    age_2 = case_when(
      age < 21 ~ "0-20",
      age < 41 ~ "21-40",
      age < 61 ~ "41-60",
      age > 60 ~ "61+"
    ),
    # Recode passive and active competence spaces
    across(ends_with(c("_read", "_write")), function(.x) ifelse(str_detect(.x, "^Yes"), 1, 0)),
    across(starts_with(c("with_", "at_", "in_", "about_")), function(.x) ifelse(str_detect(.x, "^Yes"), 1, 0)),
    across(ends_with(c("how_often")), function(.x) ifelse(str_detect(.x, "Never"), 0, 1)),
  ) %>%

  #### Passive and active competence spaces ####
  rowwise() %>%
  mutate(
    # Create passive and active competence spaces counts
    passive_spaces = sum(c_across(ends_with("_read")), na.rm = TRUE),
    active_spaces = sum(c_across(starts_with(c("with_", "at_", "in_", "about_")) | ends_with("_write")), na.rm = TRUE),
    physical_spaces = sum(c_across(space_type$new_eo[space_type$space_type == "physical"]), na.rm = TRUE),
    press_spaces = sum(c_across(space_type$new_eo[space_type$space_type == "press"]), na.rm = TRUE),
    media_spaces = sum(c_across(space_type$new_eo[space_type$space_type == "media"]), na.rm = TRUE),
    virtual_spaces = sum(c_across(space_type$new_eo[space_type$space_type == "virtual"]), na.rm = TRUE),
    other_spaces = sum(c_across(space_type$new_eo[space_type$space_type == "other"]), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    # Passive and active spaces proportion
    passive_spaces = ifelse(understand %in% c("G", "VG"), passive_spaces + 1, passive_spaces),
    # Proportion of "yes" out of 13 passive spaces
    passive_prop = passive_spaces / 13,
    # Proportion of "yes" out of 42 active spaces
    active_prop = active_spaces / 42,
    # Spaces types
    physical_tprop = physical_spaces / 23,
    press_tprop = press_spaces / 12,
    media_tprop = media_spaces / 2,
    virtual_tprop = virtual_spaces / 11,
    other_tprop = other_spaces / 2
  ) %>%
  
  #### Mean competence ####
  mutate(
    comp_counts = as.numeric(understand) +
      as.numeric(speak) +
      ifelse(read_write == "Yes", 5, 0),
    mean_comp = comp_counts / 15
  )



#### Attitudes data for MCA ####
attitudes <- emilianto %>%
  dplyr::select(educated:familiar) %>%
  drop_na()



#### Passive and active spaces long data table ####
pa_spaces <- emilianto %>%
  filter(gender != "LGBTQ+") %>%
  pivot_longer(c(passive_prop, active_prop), names_to = "spaces", values_to = "proportion") %>%
  mutate(
    spaces = recode_factor(spaces, "passive_prop" = "passive", "active_prop" = "active")
  )

#### Spaces types long data table ####
ty_spaces <- emilianto %>%
  filter(gender != "LGBTQ+") %>%
  pivot_longer(ends_with("_tprop"), names_to = "types", values_to = "proportion") %>%
  mutate(
    types = str_remove(types, "_tprop"),
    types = factor(types, levels = c("physical", "press", "media", "virtual", "other"))
  )
```

# Multiple Correspondence Analysis

This section runs the Multiple Correspondence Analysis.
The aim is to reduce the 8 qualities to a couple dimensions to simplify subsequent analyses.

```{r attitudes-mca}
attitudes_mca <- MCA(attitudes, graph = FALSE)
attitudes_dims <- attitudes_mca[["ind"]][["coord"]]
```

```{r screeplot}
fviz_screeplot(attitudes_mca, addlabels = TRUE, ylim = c(0, 15))
fviz_mca_biplot(attitudes_mca, repel = TRUE)
```

```{r mca-cor}
fviz_mca_var(attitudes_mca, choice = "mca.cor", repel = TRUE)
```

```{r mca-var}
fviz_mca_var(attitudes_mca, col.var = "black", shape.var = 15, repel = TRUE)
```

```{r mca-var-cos2}
fviz_mca_var(attitudes_mca, col.var = "cos2", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
```

```{r mca-cos2}
fviz_cos2(attitudes_mca, choice = "var", axes = 1:2)
```

```{r mca-contrib}
fviz_contrib(attitudes_mca, choice = "var", axes = 1, top = 15)
fviz_contrib(attitudes_mca, choice = "var", axes = 2, top = 15)
fviz_contrib(attitudes_mca, choice = "var", axes = 1:2, top = 15)
```

```{r mca-ind}
fviz_mca_ind(attitudes_mca, label = "none", habillage = "trustworthy")
fviz_mca_ind(attitudes_mca, label = "none", habillage = "friendly")
fviz_mca_ind(attitudes_mca, label = "none", habillage = "kind")
fviz_mca_ind(attitudes_mca, label = "none", habillage = "familiar")
```

```{r mca-dim1}
mca_dim1 <- attitudes_mca$var$coord %>%
  as_tibble(rownames = "var") %>%
  separate(var, into = c("attitude", "score"))

# Note that we are inverting Dim1 so that greater means more positive attitude.

mca_dim1_mean <- -mean(mca_dim1$`Dim 1`[mca_dim1$score == 3])

mca_dim1 %>%
  ggplot(aes(-`Dim 1`, attitude, colour = score, size = score)) +
  geom_vline(xintercept = mca_dim1_mean, linetype = "dashed") +
  geom_point() +
  scale_color_manual(values = c("#a6611a", "#dfc27d", "darkgrey", "#80cdc1", "#018571")) +
  labs(
    x = "Dimension 1 score",
    y = "Qualities",
    colour = "Likert score",
    size = "Likert score"
  )

ggsave("img/dim1.png", width = 7, height = 5)
```

```{r mca-dim2}
attitudes_mca$var$coord %>%
  as_tibble(rownames = "var") %>%
  separate(var, into = c("attitude", "score")) %>%
  ggplot(aes(`Dim 2`, attitude, colour = score, size = score)) +
  geom_point() +
  scale_color_manual(values = c("#a6611a", "#dfc27d", "darkgrey", "#80cdc1", "#018571"))
```

# Add Dimension 1

We can now create `emilianto_attitude` which includes Dimension 1 (note that this data table only contains non-NA observations in `educated:familiar`).

Note that for ease of interpretation we have reversed the values of Dimension 1 so that lower values mean a more negative attitude while more positive values mean a more positive attitude.

```{r add-dim1}
emilianto_attitude <- emilianto %>%
  drop_na(educated:familiar) %>%
  mutate(dim_1 = -attitudes_dims[, 1])
```

# Save all data tables

```{r save}
saveRDS(emilianto, "data/rds/emilianto.rds")
saveRDS(emilianto_attitude, "data/rds/emilianto_attitude.rds")
saveRDS(attitudes, "data/rds/attitudes.rds")
saveRDS(pa_spaces, "data/rds/pa_spaces.rds")
saveRDS(ty_spaces, "data/rds/ty_spaces.rds")
```

