---
title: "Analyses"
author: "Stefano Coretta"
date: "10/09/2021"
output:
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
theme_set(theme_light())
library(magrittr)
library(FactoMineR)
library(factoextra)
library(brms)
library(rnaturalearth)
library(sf)
library(tidygeocoder)
library(HH)
library(usdm)
```

# Read data

```{r}
emilianto <- readRDS("data/rds/emilianto.rds")
emilianto_attitude <- readRDS("data/rds/emilianto_attitude.rds")
attitudes <- readRDS("data/rds/attitudes.rds")
pa_spaces <- readRDS("data/rds/pa_spaces.rds")
ty_spaces <- readRDS("data/rds/ty_spaces.rds")
```


# Checks

```{r vif}
eea <- emilianto_attitude %>% dplyr::select(dim_1, passive_prop, active_prop)
vif(as.data.frame(eea))
cor(emilianto_attitude$passive_prop, emilianto_attitude$active_prop, method = "spearman")

eea <- emilianto_attitude %>% dplyr::select(dim_1, spaces_prop)
vif(as.data.frame(eea))
cor(emilianto_attitude$dim_1, emilianto_attitude$spaces_prop, method = "spearman")
```

Since `passive_prop` and `active_prop` have a somewhat high correlation (about 0.7), we decided to merge active and passive spaces into a single measure of spaces proportion `spaces_prop`.

The correlation between `spaces_prop` and `dim_1` is 0.26. The VIF is very close to 1. We believe it is safe to include both variables in the model.

# Modelling

## Understand

```{r understand-bm-prior}
get_prior(
  understand ~
    language * dim_1,
  data = emilianto_attitude,
  family = cumulative()
)

understand_priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b)
)

understand_pchek <- brm(
  understand ~
    language * dim_1,
  data = emilianto_attitude,
  family = cumulative(),
  prior = understand_priors,
  sample_prior = "only",
  backend = "cmdstanr",
  cores = 4,
  file = "./data/rds/understand_pchek"
)

conditional_effects(understand_pchek, effects = "dim_1", conditions = make_conditions(emilianto_attitude, "language"), categorical = TRUE)
```


```{r understand-bm}
understand_bm <- brm(
  understand ~
    language * dim_1,
  data = emilianto_attitude,
  family = cumulative(),
  prior = understand_priors,
  backend = "cmdstanr",
  cores = 4,
  threads = threading(2),
  file = "./data/rds/understand_bm"
)
```

```{r understand-bm-sum}
understand_bm
summary(understand_bm, prob = 0.7)
```


```{r understand-bm-cond}
ubc <- conditional_effects(understand_bm, effects = "dim_1", categorical = TRUE, conditions = make_conditions(emilianto_attitude, "language"), plot = FALSE)
```

```{r understand-bm-cond-plot}
plot(ubc, plot = FALSE)[[1]] +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_brewer(type = "div", palette = "PRGn") +
  scale_color_brewer(type = "div", palette = "PRGn") +
  labs(
    x = "Attitude (MCA Dimension 1)",
    y = "Probability",
    caption = "Attidude is operationalised as the scores of Dimension 1 from the Multiple Correspondence Analysis.\nGreater values correspond to a more positive attitude.",
    colour = "Understand\ncompetence",
    fill = "Understand\ncompetence"
  )

ggsave("img/understand-brm.png", width = 7, height = 5)
```


## Speak

```{r speak-bm}
speak_bm <- brm(
  speak ~
    language * dim_1,
  data = emilianto_attitude,
  family = cumulative(),
  # Same priors as understand_bm
  prior = understand_priors,
  backend = "cmdstanr",
  cores = 4,
  threads = threading(2),
  file = "./data/rds/speak_bm"
)
```

```{r speak-bm-sum}
speak_bm
summary(speak_bm, prob = 0.4)
```


```{r speak-bm-cond}
sbc <- conditional_effects(speak_bm, effects = "dim_1", categorical = TRUE, conditions = make_conditions(speak_bm, "language"))
```

```{r speak-bm-cond-plot}
plot(sbc, plot = FALSE)[[1]] +
  scale_fill_brewer(type = "div", palette = "PRGn") +
  scale_color_brewer(type = "div", palette = "PRGn") +
  labs(
    x = "Attitude (MCA Dimension 1)",
    y = "Probability",
    caption = "Attidude is operationalised as the scores of Dimension 1 from the Multiple Correspondence Analysis.\nGreater values correspond to a more positive attitude.",
    colour = "Speaking\ncompetence",
    fill = "Speaking\ncompetence"
  )

ggsave("img/speak-brm.png", width = 7, height = 5)
```


## Read and write

```{r rw-bm}
rw_bm <- brm(
  read_write ~
    dim_1,
  data = emilianto_attitude %>% filter(language == "Emilian"),
  family = bernoulli(),
  # Same priors as understand_bm
  prior = understand_priors,
  backend = "cmdstanr",
  cores = 4,
  threads = threading(2),
  file = "./data/rds/rw_bm"
)
```

```{r rw-bm-sum}
rw_bm
```


```{r rw-bm-cond}
conditional_effects(rw_bm, effects = "dim_1")
```

## Spaces

```{r spaces-bm}
a_spaces <- pa_spaces %>%
  filter(spaces == "active") %>%
  mutate(
    speak_bi = ifelse(speak %in% c("V", "VG"), "high", "low"),
    speak_bi = factor(speak_bi, levels = c("low", "high")),
    proportion_10 = proportion * 10
  )

spaces_bm <- brm(
  speak_bi ~
    language * proportion_10,
  data = a_spaces,
  family = bernoulli(),
  # Same priors as understand_bm
  prior = understand_priors,
  backend = "cmdstanr",
  cores = 4,
  threads = threading(2),
  file = "./data/rds/spaces_bm"
)
```

```{r spaces-bm-sum}
spaces_bm
```

```{r spaces-bm-cond}
spbc <- conditional_effects(spaces_bm, "proportion_10:language")
```

```{r spaces-bm-cond-plot}
plot(spbc, plot = FALSE)[[1]] +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_brewer(type = "qual") +
  scale_fill_brewer(type = "qual") +
  labs(
    x = "Spaces (score)",
    y = "Speaking competence (low vs high)",
    caption = str_wrap("Speaking competence was binned (low = NO/AL/50:50, high = G/VG.")
  )

ggsave("img/spaces-brm.png", width = 7, height = 5)
```


## Mean competence

```{r meanc}
emilianto_attitude %>%
  ggplot(aes(mean_comp)) +
  geom_density() +
  geom_rug()
```

```{r meanc-jitter}
emilianto_attitude %>%
  ggplot(aes(language, mean_comp)) +
  geom_jitter(width = 0.2, alpha = 0.5, height = 0) +
  expand_limits(y = 0)
```   

```{r mean-jitter-dim}
emilianto_attitude %>%
  ggplot(aes(language, mean_comp, colour = dim_1)) +
  geom_jitter(width = 0.2, alpha = 0.5, height = 0) +
  expand_limits(y = 0)
```

```{r meanc-dim1}
emilianto_attitude %>%
  ggplot(aes(dim_1, mean_comp)) +
  geom_point() +
  facet_grid(~ language)
```

```{r meanc-bm}
mean_comp_bm <- brm(
  mean_comp ~
    language *
    dim_1,
  data = emilianto_attitude,
  family = zero_one_inflated_beta(),
  backend = "cmdstanr",
  cores = 4,
  file = "./data/rds/mean_comp_bm"
)

mean_comp_bm
```

```{r meanc-bm-cond}
conditional_effects(mean_comp_bm, effects = "dim_1:language")
```


## Rural vs urban

### Understand

```{r understand-rur}
understand_rur <- emil_rur_clean %>%
  count(understand, ru_ur) %>%
  group_by(ru_ur) %>%
  mutate(
    ru_ur_prop = n / sum(n)
  ) %>%
  ungroup() %>%
  mutate(
    understand = factor(understand, ordered = FALSE),
    ru_ur = factor(ru_ur, levels = c("rural", "urban"))
  )
```

```{r understand-rur-priors}
get_prior(
  n ~
    understand * ru_ur,
  data = understand_rur,
  family = Beta()
)
```

```{r understand-rur-bm}
understand_rur_priors <- c(
  prior(normal(0, 3), class = Intercept),
  prior(normal(0, 1), class = b)
)

understand_rur_bm <- brm(
  ru_ur_prop ~
    understand * ru_ur,
  data = understand_rur,
  family = Beta(),
  prior = understand_rur_priors,
  backend = "cmdstanr",
  cores = 4,
  threads = threading(2),
  file = "./data/rds/understand_rur_bm"
)

understand_rur_bm
```

```{r understand-rur-cond}
conditional_effects(understand_rur_bm, "understand:ru_ur")
```

### Speak

```{r speak-rur}
speak_rur <- emil_rur_clean %>%
  count(speak, ru_ur) %>%
  group_by(ru_ur) %>%
  mutate(
    ru_ur_prop = n / sum(n)
  ) %>%
  ungroup() %>%
  mutate(
    speak = factor(speak, ordered = FALSE),
    ru_ur = factor(ru_ur, levels = c("rural", "urban"))
  )
```


```{r speak-rur-bm}
speak_rur_bm <- brm(
  ru_ur_prop ~
    speak * ru_ur,
  data = speak_rur,
  family = Beta(),
  prior = understand_rur_priors,
  backend = "cmdstanr",
  cores = 4,
  file = "./data/rds/speak_rur_bm"
)

speak_rur_bm
```

```{r speak-rur-cond}
conditional_effects(speak_rur_bm, "speak:ru_ur")
```

### Read and write

```{r rw-rur}
rw_rur <- emil_rur_clean %>%
  count(read_write, ru_ur) %>%
  group_by(ru_ur) %>%
  mutate(
    ru_ur_prop = n / sum(n)
  ) %>%
  ungroup() %>%
  mutate(
    ru_ur = factor(ru_ur, levels = c("rural", "urban"))
  )
```


```{r rw-rur-bm}
rw_rur_bm <- brm(
  ru_ur_prop ~
    read_write * ru_ur,
  data = rw_rur,
  family = Beta(),
  prior = understand_rur_priors,
  backend = "cmdstanr",
  cores = 4,
  file = "./data/rds/rw_rur_bm"
)

rw_rur_bm
```

```{r rw-rur-cond}
conditional_effects(rw_rur_bm, "read_write:ru_ur")
```

